---
title: "Lab 3 - Geocoding in R using Census Geocoder"
author: "Julia Kay Wolf, Ph.D.*"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_document:
    df_print: paged
    fig_height: 7
    fig_width: 7
    toc: yes
    toc_float: yes
---
*These are updates. Original version from Corey Sparks, Ph.D. can be found [here](https://rpubs.com/corey_sparks/576976).  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=F}
install.packages("censusxy")
```

# Setup
In this example, we will geocode addresses using the Census Bureau's batch geocoding service. The work will be done by the `censusxy` [package](https://cran.r-project.org/web/packages/censusxy/vignettes/censusxy.html). 

# Read in the address data
We will use the from the CSV created in the [Lab 3 exercise](https://github.com/jkwolf21/2023-DEM-5093-7093/blob/main/Code/DEM%205093_7093%20Lab%203_Spring2023.pdf).

```{r}
library(readr)
library(data.table)
addr<-read.csv("https://raw.githubusercontent.com/jkwolf21/2023-DEM-5093-7093/main/Code/wic_west_side.csv") #Note this is the RAW data* 
head(addr)
```
* See this [stack overflow discussion](https://stackoverflow.com/questions/6299220/access-a-url-and-read-data-with-r) about *raw* vs. *display* data from GitHub.

```{r}
addr<-addr[c(6, 12:14)]
names(addr)<-c("street", "city", "st", "zip")
head(addr)
library(censusxy)
results<-cxy_geocode(addr,
                     street = "street",
                     city = "city",
                     state ="st",
                     zip = "zip",
                     class="sf",
                     output = "simple")
```


# Basic interactive map of the points
```{r}
library(mapview)
mapview(results, layer.name="WIC Services")
```

# Save the results if you want
We can write the results out to a shapefile now

```{r, eval=F}
library(sf)
st_write(results,dsn="C:/Users/.../GIS Class/Lab 3", layer="westside_wic", driver = "ESRI Shapefile",delete_layer = T, append=T)
```
