---
title: "DEM 5093/7093 Lab 2 using R - Map projections and transformations"
author: "Julia Kay Wolf, Ph.D.*"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_document:
   toc: yes
   toc_float: yes
always_allow_html: yes
---
***
*These are updates. Original version from Corey Sparks, Ph.D. can be found [here](https://github.com/coreysparks/DEM7093/blob/main/code/Lab%202%20in%20R.Rmd).  

This lab complements the Lab 2 exercise using QGIS. 

Here, we use `tidycensus` to read some tract data, learn its projection information, transform it to a new coordinate system, and measure some distance between features.

***

### Call Packages from Library
```{r}
library(tidycensus)
library(sf)
library(dplyr)
```
**Note: you can run `quietly=T` after any `library` command to remove the messages in the white boxes with ## in your output, e.g., `library(sf, quietly=T)`** 


### Read in Bexar county tracts

*(Seem familiar? We did this chunk below before in Lab 1/Homework 2.)*

```{r, results="hide"}
sa_acs<-get_acs(geography = "tract",
                state="TX",
                county = c("Bexar"),
                year = 2019,
                variables=c( "DP05_0001E", 
                            "DP03_0119PE") ,
                geometry = T, output = "wide")
#create a county FIPS code - 5 digit
sa_acs$county<-substr(sa_acs$GEOID, 1, 5)
#rename variables and filter missing cases
sa_acs2<-sa_acs%>%
  mutate(totpop= DP05_0001E, ppov=DP03_0119PE) %>%
#  st_transform(crs = 102740)%>%
  na.omit()
```


### Find Coordinate System of Current Map

**See [Coordinate Systems, Projections, and Transformations](https://pro.arcgis.com/en/pro-app/latest/help/mapping/properties/coordinate-systems-and-projections.htm) from ArcGIS Pro/ESRI for more information about these concepts.**
```{r}
st_crs(sa_acs2)
```
We see these tracts are in a Geographic Coordinate System (GCS) called North American Datum 1983 (NAD83).

### Create Basic Map
```{r}
library(tmap)
library(tmaptools)
tm_shape(sa_acs2)+
  tm_polygons("ppov", title="% in Poverty",
              palette="Blues",
              style="quantile",
              n=5 )+
  tm_format("World",
            main.title="San Antonio Poverty Estimates (2019) - Quintile Breaks",
            main.title.position=c('center','top'),
            main.title.size=1.5,
            title="Author: Julia Kay Wolf, Ph.D. \nSource: ACS 2019",
            legend.title.size=1.7,
            legend.outside=T,
            legend.text.size=1.2)+
  tm_scale_bar(position = c("left","bottom"))+
  tm_compass()
```

Click [here](https://www.rdocumentation.org/packages/tmap/versions/3.3-3/topics/tm_layout) for more `tmap` aesthetic features.  

Click [here](https://www.statology.org/percentile-vs-quartile-vs-quantile/) for a quick discussion on quantile vs. quintile. 


### Re-Project Map into "South Central Texas" Projection

*(Remember 2278 from the QGIS portion of Lab 2?)*

```{r}
new_sa<-st_transform(sa_acs2, crs = 2278)
#Extract two tracts
twtr<-new_sa%>%
  filter(GEOID %in% c(48029181820, 48029110600))
# get centroid coordinates for two tracts (these two tracts are where UTSA Main and Downtown Campuses are)
tr_co<-st_centroid(twtr)
head(tr_co)
```


## Measure Feet Apart
```{r}
st_distance(tr_co)
64041.12/5280 #To get feet into miles
```
*(Remember in QGIS we got 12.13 miles?)*


# Using QGIS within R
This is another way to do the above task, by running a QGIS algorithm within R using the `qgisprocess` package. 

*NOTE: The `qgisprocess` will not install on its own with this 4.2.2 version of R.*


See the vignette [here](https://github.com/paleolimbot/qgisprocess) for more on what this package is and some examples.
See [here](https://paleolimbot.github.io/qgisprocess/) to explain the following chunk for installing it.

```{r, eval=F}
install.packages("remotes")
remotes::install_github("paleolimbot/qgisprocess") #Select 1 for ALL; next there will be a pop-up box, select Yes.
```


```{r}
library(qgisprocess) #load the package
qgis_configure() #set up QGIS - find the executable
# qgis_algorithms() lists all the available routines in QGIS
head(qgis_algorithms())
```

We can use `grep` to search for specific terms in the algorithms. 

```{r}
algs<-qgis_algorithms()
algs[grep(x = algs$algorithm, "distance"),"algorithm"]
qgis_show_help("qgis:distancematrix")
```

```{r}
out = qgis_run_algorithm(alg = "qgis:distancematrix",
               INPUT = tr_co[1,],
               INPUT_FIELD = "GEOID", 
               TARGET = tr_co[2,],
               TARGET_FIELD = "GEOID",
               MATRIX_TYPE = 0, 
               NEAREST_POINTS = 1)
output_sf <- sf::read_sf(qgis_output(out, "OUTPUT"))
output_sf$Distance
64041.12/5280 #To get feet into miles
```
NOTE: `tr_co[1,]` means select the first row from `tr_co` which is one of our two tracts. `tr_co[2,]` means select the 2nd row, which is our other tract.

*(See it's the same 12.13 mile distance?)* 
